Online Multiclass Linear Classification with Bandit Feedback

### 2018-11-27 arxiv-version/main.tex
Improve intro, add a reference

We design new algorithms for the problem of \textsc{Online Multiclass
Classification with Bandit Feedback}. In this problem, in each round, the
algorithm receives a feature vector and the algorithm has to predict one of $K$
possible classes. Immediately after the algorithm makes its prediction, it
receives feedback whether the prediction was correct or not. However, in
contrast with the standard, so called full information setting, the algorithm
does \emph{not} receive the correct label. The feedback is only binary. To allow
for meaningful randomized strategies, we assume that the correct label in a
particular round is chosen prior to algorithm's prediction in that
round.\footnote{The reader can assume throughout the paper a simpler model of
the so called \emph{oblivious adversary} where the whole sequence of labeled
examples $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T)$ is chosen before the first
round of the game.} The protocol of the problem is stated below.

@@

The bandit feedback setting, however, is much more challenging. To understand
the differences, we dissect the notion of linear separability in the multiclass
setting. Beside the standard notion of linear separability, we define also
\emph{strong linear separability}. For binary classification, these two
definitions are identical. However, they differ for multiclass classification.
For multiclass classification with $K$ classes the standard notion of linear
separability means that the examples from each class lie in an intersection of
$K-1$ halfspaces and the examples outside of the class lie in the complement of
the intersection of the halfspaces. Strong linear separability means that
examples from each class is separated from the remaining examples by a
\emph{single} hyperplane. We state the definitions precisely in
Section~\ref{section:notions-of-linear-separability}.

@@

For the standard notion of linear separability, we propose three algorithms with
various time-complexity-versus-mistake-bound trade-offs. Two of these algorithms
map the examples into a high-dimensional inner product space induced by a
kernel. In this space the examples become \emph{strongly} linearly separable.
This allows us to run a kernelized version of the algorithm for the strongly
linearly separable case.

The two kernel-based algorithms differ by the choice of the kernel. The crux of
the problem is the analysis of the approximation properties of these kernels and
the way margin in original space under standard linear separability notion gets
transformed into a margin in the high-dimensional space under the strong
separability notion. This problem is related to the problem of learning
intersection of halfspaces and has been studied previously by
\cite{Klivans-Servedio-2008}.

The third algorithm is based on the obvious idea that two points that are close
enough must have the same label. Here, close enough means that they cannot be
separated from each other by a linear separator with a margin.

All three algorithms run in time that is polynomial in the dimension of the
feature vectors, the number of classes and the number of rounds. At the same
time, all three algorithms make number of mistakes that does \emph{not} depend
on the number of rounds. Instead, their number of mistakes depends only on the
margin, number of classes, dimension of the feature vectors, and the radius of
the ball in which the data lie.

Finally, we study two questions related to the computational and
information-theoretic hardness of the problem. Any algorithm for the bandit
setting collects information in the form of so called \emph{strongly labeled}
and \emph{weakly labeled} examples. Strongly-labeled examples are those for
which we know the class label. Weakly labeled example is an example for which we
know that class label can be anything except for a particular one class. First,
we show the offline problem of determining if a linear multiclass separator
exists for a set of strongly- and weakly-labeled examples is NP-complete. This
result is a simple consequence of the fact that learning intersection of
halfspaces is NP-complete~\citep{Blum-Rivest-1993}. Second, we show that any
algorithm that uses only strongly-labeled examples and ignores weakly labeled
examples makes at least $\Omega(???)$ mistakes.

@@

\cite{Abernethy-Rakhlin-2009}

\cite{Hazan-Kale-2011}

\cite{Foster-Krishnamurthy-2018}

>>>>

We design new algorithms for the problem of \textsc{Online Multiclass
Classification with Bandit Feedback}. In this problem, in each round, the
algorithm receives a feature vector and the algorithm has to predict one of $K$
possible classes. Immediately after the algorithm makes its prediction, it
receives feedback whether the prediction was correct or not. However, in
contrast with the standard, so called full information setting, the algorithm
does \emph{not} receive the correct label. The feedback is only binary. To allow
for meaningful use of randomization, we assume that the correct label in a
particular round is chosen prior to algorithm's prediction in that
round.\footnote{The reader can assume throughout the paper a simpler model of
the so called \emph{oblivious adversary} where the whole sequence of labeled
examples $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T)$ is chosen before the first
round of the game.} The protocol of the problem is stated below.

@@

The bandit feedback setting, however, is much more challenging. To understand
the differences, we dissect the notion of linear separability in the multiclass
setting. Beside the standard notion of linear separability, we define also
\emph{strong linear separability}. For binary classification, these two
definitions are identical. However, they differ for multiclass classification.
For multiclass classification with $K$ classes the standard notion of linear
separability means that the examples from each class lie in an intersection of
$K-1$ halfspaces and the examples outside of the class lie in the complement of
the intersection of the halfspaces. Strong linear separability means that
examples from each class are separated from the remaining examples by a
\emph{single} hyperplane. We state the definitions precisely in
Section~\ref{section:notions-of-linear-separability}.

@@

For the standard notion of linear separability, we propose several algorithms
with various time-complexity-versus-mistake-bound trade-offs. First class of
algorithms map the examples into a high-dimensional inner product space. In this
space the examples become \emph{strongly} linearly separable. This allows us to
run the algorithm for the strongly linearly separable case. All of these
algorithms can be kernelized. (Some of them have to be kernalized since the
underlying inner product space is infinite dimensional.) Kernelization does not
change the input-output behavior of the algorithm, it only affects its time
complexity.

These algorithms differ by choice of the mapping or, equivalently, choice of the
kernel. The crux of the problem is the analysis of the approximation properties
of these mappings/kernels and the way margin in original space under standard
linear separability notion gets transformed into a margin in the
high-dimensional space under the strong separability notion. This problem is
related to the problem of learning intersection of halfspaces and has been
studied previously by \cite{Klivans-Servedio-2008}.

Finally, we propose a very different algorithm that is based on the obvious idea
that two points that are close enough must have the same label. Here, close
enough means that they cannot be separated from each other by a linear separator
with a margin.

All the algorithms run in time that is polynomial in the dimension of the
feature vectors, the number of classes and the number of rounds. At the same
time, all the algorithms make number of mistakes that does \emph{not} depend
on the number of rounds. Instead, their number of mistakes depends only on the
margin, number of classes, dimension of the feature vectors, and the radius of
the ball in which the data lie.

Finally, we study two questions related to the computational and
information-theoretic hardness of the problem. Any algorithm for the bandit
setting collects information in the form of so called \emph{strongly labeled}
and \emph{weakly labeled} examples. Strongly-labeled examples are those for
which we know the class label. Weakly labeled example is an example for which we
know that class label can be anything except for a particular one class. First,
we show the offline problem of finding a linear multiclass separator consistent
with the strongly- and weakly-labeled examples is NP-hard. This result is a
simple consequence of the fact that finding an intersection of two halfspaces
consistent with binary-labeled examples is NP-hard~\citep{Blum-Rivest-1993}.
Second, we show that any algorithm that uses only strongly-labeled examples and
ignores weakly labeled examples makes at least $\Omega(???)$ mistakes.

@@

\begin{itemize}
\item \cite{Abernethy-Rakhlin-2009}

\item \cite{Hazan-Kale-2011}

\item \cite{Beygelzimer-Orabona-Zhang-2017}

\item \cite{Foster-Krishnamurthy-2018}
\end{itemize}
